---
title: "Co-regulation across populations (MESA data)"
---

```{r setup}
rm(list = ls())
library(data.table)
options(datatable.fread.datatable = F)
library(SilverStandardPerformance)
library(dplyr)
library(ggplot2)
```


```{r func}
fast_corr = function(x) {
  x = as.matrix(x)
  mean_ = apply(x, 1, mean)
  sd_ = apply(x, 1, mysd)
  x_norm = (x - mean_) / sd_
  cor_ = x_norm %*% t(x_norm) / ncol(x_norm)
  cor_
}

inverse_normalize = function(x, offset = 1) {
  g = rank(x, ties.method = 'average') / (length(x) + offset)
  o = qnorm(g)
  return(o)
}

mysd = function(x) {
  sqrt(mean((x - mean(x)) ^ 2))
}

annot_color = function(pop, color_panel) {
  e = data.frame(pop = pop)
  f = left_join(e, color_panel, by = 'pop')
  f$color
}

myimage = function(cor_mat, indiv_color_panel = NULL) {
  cor_mat_melt = melt(cor_mat)
  p = cor_mat_melt %>% ggplot() + 
    geom_raster(aes(x = Var1, y = Var2, fill = value)) + scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red')
  p = p + theme(
    plot.title = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.line = element_blank()
  )
  if(!is.null(indiv_color_panel)) {
    p = p + 
    geom_point(aes(x = Var1, y = -2, color = Var1), show.legend = F, shape = 15, size = 1) + 
    geom_point(aes(x = -1, y = Var2, color = Var2), show.legend = F, shape = 15, size = 1) + 
    scale_color_manual(values = indiv_color_panel)
  }
  p
}

get_partition = function(indiv_list, test_size) {
  test_idx = sample(1 : length(indiv_list), size = test_size, replace = FALSE)
  test_ind = (1 : length(indiv_list)) %in% test_idx
  list(reference = indiv_list[!test_ind], test = indiv_list[test_ind])
}

mat_dist = function(m1, m2) {
  sqrt(sum((m1 - m2) ^ 2))
}
```

```{r param}
candidate_colors = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Load data

From Box [link](https://uchicago.box.com/s/gcyzzk82y90b6la3pm7fbdyc4tmic24i).
Here I used RPKM which is normalized by library size and normalized by gene length (this does not matter since we do inverse normalization gene by gene).
First **filter** out individuals with **analysis freeze = FALSE**. 
Secondly, **filter** out individuals with **X3p_bias** missing (I have no idea what that means).   
Thridly, **limit** to PBMC samples.   

```{r}
data("gene_annotation_gencode_v26_hg38")
head(gene_annotation_gencode_v26_hg38$gene_annotation)

indiv_meta = read.csv('~/Desktop/tmp/TOPMed_MESA/TOPMed_MESA_RNAseq_sample_attributes_freeze.csv')
indiv_meta = indiv_meta %>% filter(analysis_freeze, race != '') %>% filter(!is.na(X3p_bias)) %>% filter(sample_type == 'PBMC')
indiv_meta = indiv_meta[order(indiv_meta$race), ]

df = fread('zcat < ~/Desktop/tmp/TOPMed_MESA/TOPMed_MESA_RNAseq_Pilot_RNASeQCv2.0.0.gene_rpkm.gct.gz', header = T)
df_indiv = colnames(df)[c(-1, -2)]
df_cleaned = df[, c(T, T, df_indiv %in% indiv_meta$tor_id)]
rownames(df_cleaned) = df$Name
```

# Some further QC's

Limit to protein coding and/or lincRNA.
Limit to genes with non-zero observations in all individuals.
**For each population**, do inverse normalization gene by gene to obtain the expression matrix.
Limit to top 1000 highly expression genes (expression is measured as the median of RPKM across all individuals).

```{r}
df_protein_or_linc = df_cleaned %>% filter(Description %in% gene_annotation_gencode_v26_hg38$gene_annotation$gene_name[gene_annotation_gencode_v26_hg38$gene_annotation$gene_type %in% c('protein_coding', 'lincRNA')])

n_zeros = apply(df_protein_or_linc[, c(-1, -2)], 1, function(x) {
  sum(x == 0)
})

rownames(df_protein_or_linc) = df_protein_or_linc$Name
df_protein_or_linc_non_zero = df_protein_or_linc[n_zeros == 0, ]
expr_level = apply(df_protein_or_linc_non_zero[, c(-1, -2)], 1, function(x) {
  median(x)
})
top_1000_expr_level_ind = rank(-expr_level) %in% 1 : 1000
df_protein_or_linc_non_zero = df_protein_or_linc_non_zero[top_1000_expr_level_ind, ]

tmp = df_protein_or_linc_non_zero[, 3 : ncol(df_protein_or_linc_non_zero)]
tmp = tmp[, match(indiv_meta$tor_id, colnames(tmp))]
df_protein_or_linc_non_zero = cbind(df_protein_or_linc_non_zero[, 1:2], tmp)


# some meta operations
indiv_meta %>% group_by(race) %>% summarize(n = n())

color_panel = data.frame(pop = unique(indiv_meta$race), color = candidate_colors[1 : length(unique(indiv_meta$race))])
indiv_color_panel = indiv_meta %>% left_join(color_panel, by = c('race' = 'pop')) %>% pull(color)
names(indiv_color_panel) = indiv_meta$tor_id
```

Perform normalization for each populations.

```{r normalization}
# among all
normalized_mat = t(apply(df_protein_or_linc_non_zero[, c(-1, -2)], 1, inverse_normalize))

# for each population
mat = df_protein_or_linc_non_zero[, c(-1, -2)]
normalized_mat_by_pop = matrix(NA, ncol = ncol(normalized_mat), nrow = nrow(normalized_mat))
start = 1
starts = c()
names_ = c()
for(rr in unique(indiv_meta$race)) {
  starts = c(starts, start)
  message(rr, ' start = ', start)
  indiv_here = indiv_meta$tor_id[indiv_meta$race == rr]
  mat_sub = mat[, colnames(mat) %in% indiv_here]
  normalized_mat_by_pop[, start : (length(indiv_here) + start - 1)] = t(apply(mat_sub, 1, inverse_normalize))
  start = length(indiv_here) + start 
  names_ = c(names_, colnames(mat_sub))
}
colnames(normalized_mat_by_pop) = names_
rownames(normalized_mat_by_pop) = rownames(normalized_mat)
```

# Between individual correlation

```{r}
cor_indiv = fast_corr(t(as.matrix(normalized_mat)))
cor_indiv_by_pop = fast_corr(t(as.matrix(normalized_mat_by_pop)))
```

Visualization.

```{r vis}
# image Asian 
myimage(cor_indiv[1:90, 1:90], indiv_color_panel)

myimage(cor_indiv_by_pop[1:90, 1:90], indiv_color_panel)
# each taking 40
indexes = c()
for(i in starts) {
  indexes = c(indexes, i : (i + 40 - 1))
}
myimage(cor_indiv[indexes, indexes], indiv_color_panel)
myimage(cor_indiv_by_pop[indexes, indexes], indiv_color_panel)
```

# Between gene correlation (using Black as an example)

As a test case, I calculate the gene correlation matrix within Black (using inverse normalized expression within each population).
Here we are interested in measuring the similarity between Black's gene correlation and White's gene correlation. 
To account for the sample standard error, we partition the White samples into two parts: 1) **test part**: with equal size to Black samples; 2) **reference part**: the rest (it should be larger part).
For the first partition, we calculate the distance between Black and **reference part**. And for the other partitions, we calcualte the distance between **test part** and **reference part**. Here the distance is the test statistic. 

```{r cor gene}
set.seed(1)
### The following notes are deprecated!
### IMPORTANT!!!
# for fast testing, I randomly select 500 genes. 
# it is just for testing code!
normalized_mat_by_pop_debug = normalized_mat_by_pop # [sample(nrow(normalized_mat_by_pop), size = 500), ]
### END

num_partitions = 100

test_pop = 'Black'
test_indiv_here = indiv_meta$tor_id[indiv_meta$race == test_pop]
expr_test = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% test_indiv_here]
mat_test = fast_corr(expr_test)

ref_pop = 'White'
ref_indiv_here = indiv_meta$tor_id[indiv_meta$race == ref_pop]
partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
mat_ref = fast_corr(expr_ref)

test_u = mat_dist(mat_test, mat_ref)

test_null = c()
for(i in 1 : num_partitions) {
  partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
  expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
  mat_ref = fast_corr(expr_ref)
  expr_tref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$test]
  mat_tref = fast_corr(expr_tref)
  test_null = c(test_null, mat_dist(mat_tref, mat_ref))
}

# 
# myimage(mat_test[1:100, 1:100])
# myimage(mat_ref[1:100, 1:100])
hist(test_null, xlim = c(min(c(test_null, test_u)), max(c(test_null, test_u)))); abline(v = test_u, col = 'red')
```

# How about Hispanic?

```{r cor gene hispanic}
set.seed(1)
### The following notes are deprecated!
### IMPORTANT!!!
# for fast testing, I randomly select 500 genes. 
# it is just for testing code!
normalized_mat_by_pop_debug = normalized_mat_by_pop # [sample(nrow(normalized_mat_by_pop), size = 500), ]
### END

num_partitions = 100

test_pop = 'Hispanic'
test_indiv_here = indiv_meta$tor_id[indiv_meta$race == test_pop]
expr_test = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% test_indiv_here]
mat_test = fast_corr(expr_test)

ref_pop = 'White'
ref_indiv_here = indiv_meta$tor_id[indiv_meta$race == ref_pop]
partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
mat_ref = fast_corr(expr_ref)

test_u = mat_dist(mat_test, mat_ref)

test_null = c()
for(i in 1 : num_partitions) {
  partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
  expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
  mat_ref = fast_corr(expr_ref)
  expr_tref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$test]
  mat_tref = fast_corr(expr_tref)
  test_null = c(test_null, mat_dist(mat_tref, mat_ref))
}

# 
# myimage(mat_test[1:100, 1:100])
# myimage(mat_ref[1:100, 1:100])
hist(test_null, xlim = c(min(c(test_null, test_u)), max(c(test_null, test_u)))); abline(v = test_u, col = 'red')
```

# How about Asian?

```{r cor gene asian}
set.seed(1)
### The following notes are deprecated!
### IMPORTANT!!!
# for fast testing, I randomly select 500 genes. 
# it is just for testing code!
normalized_mat_by_pop_debug = normalized_mat_by_pop # [sample(nrow(normalized_mat_by_pop), size = 500), ]
### END

num_partitions = 100

test_pop = 'Asian'
test_indiv_here = indiv_meta$tor_id[indiv_meta$race == test_pop]
expr_test = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% test_indiv_here]
mat_test = fast_corr(expr_test)

ref_pop = 'White'
ref_indiv_here = indiv_meta$tor_id[indiv_meta$race == ref_pop]
partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
mat_ref = fast_corr(expr_ref)

test_u = mat_dist(mat_test, mat_ref)

test_null = c()
for(i in 1 : num_partitions) {
  partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
  expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
  mat_ref = fast_corr(expr_ref)
  expr_tref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$test]
  mat_tref = fast_corr(expr_tref)
  test_null = c(test_null, mat_dist(mat_tref, mat_ref))
}

# 
# myimage(mat_test[1:100, 1:100])
# myimage(mat_ref[1:100, 1:100])
hist(test_null, xlim = c(min(c(test_null, test_u)), max(c(test_null, test_u)))); abline(v = test_u, col = 'red')
```

# How about Black with downsample to 40?

```{r cor gene black}
set.seed(1)
### The following notes are deprecated!
### IMPORTANT!!!
# for fast testing, I randomly select 500 genes. 
# it is just for testing code!
normalized_mat_by_pop_debug = normalized_mat_by_pop # [sample(nrow(normalized_mat_by_pop), size = 500), ]
### END

num_partitions = 100
downsample_size = 40

test_pop = 'Black'
test_indiv_here = indiv_meta$tor_id[indiv_meta$race == test_pop]
test_indiv_here = test_indiv_here[sample(length(test_indiv_here), size = downsample_size)]
expr_test = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% test_indiv_here]
mat_test = fast_corr(expr_test)

ref_pop = 'White'
ref_indiv_here = indiv_meta$tor_id[indiv_meta$race == ref_pop]
partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
mat_ref = fast_corr(expr_ref)

test_u = mat_dist(mat_test, mat_ref)

test_null = c()
for(i in 1 : num_partitions) {
  partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
  expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
  mat_ref = fast_corr(expr_ref)
  expr_tref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$test]
  mat_tref = fast_corr(expr_tref)
  test_null = c(test_null, mat_dist(mat_tref, mat_ref))
}

# 
# myimage(mat_test[1:100, 1:100])
# myimage(mat_ref[1:100, 1:100])
hist(test_null, xlim = c(min(c(test_null, test_u)), max(c(test_null, test_u)))); abline(v = test_u, col = 'red')
```

# How about Hispanic with downsample to 40?

```{r cor gene hispanic 40}
set.seed(1)
### The following notes are deprecated!
### IMPORTANT!!!
# for fast testing, I randomly select 500 genes. 
# it is just for testing code!
normalized_mat_by_pop_debug = normalized_mat_by_pop # [sample(nrow(normalized_mat_by_pop), size = 500), ]
### END

num_partitions = 100
downsample_size = 40

test_pop = 'Hispanic'
test_indiv_here = indiv_meta$tor_id[indiv_meta$race == test_pop]
test_indiv_here = test_indiv_here[sample(length(test_indiv_here), size = downsample_size)]
expr_test = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% test_indiv_here]
mat_test = fast_corr(expr_test)

ref_pop = 'White'
ref_indiv_here = indiv_meta$tor_id[indiv_meta$race == ref_pop]
partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
mat_ref = fast_corr(expr_ref)

test_u = mat_dist(mat_test, mat_ref)

test_null = c()
for(i in 1 : num_partitions) {
  partition_here = get_partition(ref_indiv_here, test_size = length(test_indiv_here))
  expr_ref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$reference]
  mat_ref = fast_corr(expr_ref)
  expr_tref = normalized_mat_by_pop_debug[, colnames(normalized_mat_by_pop_debug) %in% partition_here$test]
  mat_tref = fast_corr(expr_tref)
  test_null = c(test_null, mat_dist(mat_tref, mat_ref))
}

# 
# myimage(mat_test[1:100, 1:100])
# myimage(mat_ref[1:100, 1:100])
hist(test_null, xlim = c(min(c(test_null, test_u)), max(c(test_null, test_u)))); abline(v = test_u, col = 'red')
```
