---
title: "First look at MESA PTRS along with PTRS with LD block pruning"
---

```{r setup}
library(dplyr)
library(data.table)
library(ggplot2)
options(stringsAsFactors = F)
parse_pop = function(x) {
  x = unlist(lapply(strsplit(x, '-'), function(y) {
    if(length(y) > 1) {
      return(paste0(y[1], '-', y[2]))
    } else {
      return(y)
    }
  }))
}
theme_set(theme_bw(base_size = 10))
source('../code/rlib_doc.R')
traits = tolower(read.csv('../external_data/martin_et_al_2019ng_table_s6.csv')$Trait)
diag_df = data.frame(trait = traits, index = 1 : 17)
```

# About 

Gene weights were obtained from S-PrediXcan run using MESA CAU model.

# Load results

## MESA PTRS

```{r load}
mesa_models = c('AFA', 'AFHI', 'CAU', 'ALL', 'HIS')
result = list()
for(m in mesa_models) {
  for(i in 1 : 17) {
    for(t in traits) {
      filename = paste0('~/Desktop/tmp/ptrs_mesa_r2/ptrs_mesa_r2_', m, '/ptrs-r2_subset', i, '_x_', t, '.txt')
      tmp = read.table(filename, header = T, sep = '\t')
      result[[length(result) + 1]] = tmp %>% mutate(subset = i, trait = t, model = m)
    }
  }
}
df_mesa = do.call(rbind, result)
df_mesa$population = parse_pop(df_mesa$population)
df_mesa$ptrs_col = factor(df_mesa$ptrs_col, levels = unique(df_mesa$ptrs_col))
```

## MESA PTRS with LD block based pruning

```{r load33}
mesa_models = c('AFHI', 'CAU', 'ALL')
result = list()
for(m in mesa_models) {
  for(i in 1 : 17) {
    for(t in traits) {
      filename = paste0('~/Desktop/tmp/ptrs_mesa_ldblock_r2/ptrs_mesa_ldblock_r2_', m, '/ptrs-r2_subset', i, '_x_', t, '.txt')
      tmp = read.table(filename, header = T, sep = '\t')
      result[[length(result) + 1]] = tmp %>% mutate(subset = i, trait = t, model = m)
    }
  }
}
df_mesa_ld = do.call(rbind, result)
df_mesa_ld$population = parse_pop(df_mesa_ld$population)
df_mesa_ld$ptrs_col = factor(df_mesa_ld$ptrs_col, levels = unique(df_mesa_ld$ptrs_col))
```

## GTEx V8 CTIMP whole blood PTRS 

```{r load2}
result = list()
for(i in 1 : 17) {
  for(t in traits) {
    filename = paste0('~/Desktop/tmp/ptrs_r2/ptrs-r2_subset', i, '_x_', t, '.txt')
    tmp = read.table(filename, header = T, sep = '\t')
    result[[length(result) + 1]] = tmp %>% mutate(subset = i, trait = t, model = 'gtex_v8_ctimp')
  }
}
df_gtex = do.call(rbind, result)
df_gtex$population = parse_pop(df_gtex$population)
df_gtex$ptrs_col = factor(df_gtex$ptrs_col, levels = unique(df_gtex$ptrs_col))
```

## Merge

```{r merge}
df_ptrs = rbind(df_mesa %>% select(-SSE.wo, -SSE.with), df_gtex)
df_ld = rbind(df_mesa %>% filter(model %in% mesa_models) %>% select(-SSE.wo, -SSE.with) %>% mutate(type = 'naive'), df_mesa_ld %>% filter(model %in% mesa_models) %>% select(-SSE.wo, -SSE.with) %>% mutate(type = 'ldblock'))
```

# Plot

## Overview

```{r overview, fig.height=10, fig.width=10}
df_ptrs %>% ggplot() + geom_boxplot(aes(x = ptrs_col, y = r2, color = population)) + facet_wrap(~model, ncol = 2, scales = 'free_y') + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + theme(legend.position = 'bottom')
```

## Zoom in AFHI model vs CAU model

```{r afa-vs-cau}
df_ptrs %>% filter(model %in% c('CAU', 'AFHI')) %>% select(-pval) %>% dcast(population + ptrs_col + subset + trait ~ model, value.var = 'r2') %>% filter(paste(trait, subset) %in% paste(diag_df$trait, diag_df$index)) %>% 
  ggplot() + geom_point(aes(x = CAU, y = AFHI), alpha = .3) + facet_wrap(~population, scales = 'free') + geom_abline(slope = 1, intercept = 0)
```

## Best CAU model performance

```{r best-cau}
df_cau = df_ptrs %>% filter(model == 'CAU')
best_cau = best_model_based_on_one(df_cau, pop_name = 'British-validation', model_col = 'ptrs_col', score_col = 'r2')
best_cau$perf_in_all %>% 
  filter(paste(trait, subset) %in% paste(diag_df$trait, diag_df$index)) %>% 
  ggplot() + geom_violin(aes(x = population, y = transferability)) + 
  geom_jitter(aes(x = population, y = transferability), height = 0, width = 0.2)
```

## Performance of CAU model with other prediction models

```{r analysis-function}
performance_of_best_model = function(best_model, best_pop, candidate_models) {
  df_candidate_best = candidate_models %>% filter(paste(trait, ptrs_col, subset) %in% paste(best_model$best_model$trait, best_model$best_model$best_model, best_model$best_model$subset))
  df_candidate_best = inner_join(df_candidate_best, best_model$perf_in_all %>% filter(population == best_pop) %>% select(trait, ptrs_col, subset, r2) %>% rename(best_r2 = r2), by = c('trait', 'ptrs_col', 'subset'))
  df_candidate_best_perf = df_candidate_best %>% mutate(transferability = r2 / best_r2)
  df_candidate_best_perf = inner_join(best_model$perf_in_all, df_candidate_best_perf, by = c('trait', 'population', 'subset', 'ptrs_col'), suffix = c('_best', '_cand')) %>% mutate(ratio_cand_over_best = r2_cand / r2_best)
  df_candidate_best_perf
}
```

Loop over all other prediction models.

```{r best-cau-with-others}
out_list = list()
for(m in mesa_models) {
  # if( m == 'CAU') {
  #   next
  # }
  df_tmp = performance_of_best_model(best_cau, 'British-validation', df_ptrs %>% filter(model == m))
  out_list[[length(out_list) + 1]] = df_tmp %>% mutate(other_model = m)
}
df_best_cau_all_others = do.call(rbind, out_list)
```


Performance of best CAU model with other prediction models.

```{r perf-best-cau-with-other}
df_best_cau_all_others %>% 
  filter(paste(trait, subset) %in% paste(diag_df$trait, diag_df$index)) %>% 
  ggplot() + geom_violin(aes(x = population, y = transferability_cand, color = other_model)) + 
  geom_jitter(aes(x = population, y = transferability_cand, color = other_model), position = position_dodge(width = .9))
```

Side-by-side comparing best CAU model with other prediction models

```{r side-by-side-best-cau-with-others, fig.height=6, fig.width=6}
df_best_cau_all_others %>% filter(population != 'British-validation', other_model != 'CAU') %>% 
  ggplot(aes(x = transferability_best, y = transferability_cand)) + 
  geom_point(alpha = 0.1) + 
  facet_grid(population~other_model, scales = 'free') + 
  geom_abline(slope = 1, intercept = 0) + 
  scale_x_log10() + 
  scale_y_log10()
df_best_cau_all_others %>% filter(population != 'British-validation', other_model != 'CAU') %>%
  ggplot() + geom_boxplot(aes(x = trait, y = ratio_cand_over_best), alpha = 0.3) + facet_grid(population~other_model, scales = 'free_y') + scale_y_log10() + geom_hline(yintercept = 1) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

## Best model within each

Calculate.

```{r best for each}
best_for_each = df_ptrs %>% filter(model != 'gtex_v8_ctimp') %>% group_by(trait, subset, model, population) %>% summarize(best_model = ptrs_col[which.max(r2)], max_r2 = max(r2)) %>% ungroup()
best_for_each = inner_join(best_for_each, best_cau$perf_in_all %>% filter(population == 'British-validation') %>% select(r2, trait, subset) %>% rename(reference_r2 = r2), by = c('trait', 'subset'))
best_for_each = best_for_each %>% mutate(transferability = max_r2 / reference_r2)
```

Plot transferability.

```{r plot-best-for-each}
best_for_each %>% filter(paste(trait, subset) %in% paste(diag_df$trait, diag_df$index)) %>% 
  ggplot() + geom_violin(aes(x = population, y = transferability, color = model)) + 
  geom_jitter(aes(x = population, y = transferability, color = model), position = position_dodge(width = .9))
```

Side-by-side transferability between prediction models.

```{r calc}
best_for_each_reference = best_for_each %>% filter(model == 'CAU')
best_for_each = inner_join(best_for_each, best_for_each_reference %>% select(trait, subset, population, transferability), by = c('trait', 'subset', 'population'), suffix = c('_cand', '_best')) 
best_for_each = best_for_each %>% mutate(ratio_cand_over_best = transferability_cand / transferability_best)
```

```{r plot-best-for-each2}
best_for_each %>% filter(population != 'British-validation', model != 'CAU') %>% 
  ggplot(aes(x = transferability_best, y = transferability_cand)) + 
  geom_point(alpha = 0.1) + 
  facet_grid(population~model, scales = 'free') + 
  geom_abline(slope = 1, intercept = 0) + 
  scale_x_log10() + 
  scale_y_log10()
best_for_each %>% filter(population != 'British-validation', model != 'CAU') %>%
  ggplot() + geom_boxplot(aes(x = trait, y = ratio_cand_over_best), alpha = 0.3) + facet_grid(population~model, scales = 'free_y') + scale_y_log10() + geom_hline(yintercept = 1) + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
```

## Compare with PTRS PVE

Load

```{r load-ptrs-pve}
regu = readRDS('../analysis_output/regulability_mesa_and_ctimp.rds')
df_best_cau_all_others$pop = df_best_cau_all_others$population
df_best_cau_all_others$pop[df_best_cau_all_others$pop %in% c('British-test', 'British-validation')] = 'British'
df_best_cau_all_others = inner_join(df_best_cau_all_others, regu, by = c('trait', 'pop' = 'population', 'model'))
best_for_each$pop = best_for_each$population
best_for_each$pop[best_for_each$pop %in% c('British-test', 'British-validation')] = 'British'
best_for_each = inner_join(best_for_each, regu, by = c('trait', 'pop' = 'population', 'model'))
```

Plot

```{r ptrs-pve-vs-r2-best-cau, fig.height=6, fig.width=6}
df_best_cau_all_others %>% 
  filter(paste(trait, subset) %in% paste(diag_df$trait, diag_df$index)) %>% 
  ggplot() + geom_point(aes(x = h_sq, y = r2_cand)) + 
  facet_grid(other_model~population, scales = 'free') 
  
```

```{r ptrs-pve-vs-r2-best-each, fig.height=6, fig.width=6}
best_for_each %>% 
  filter(paste(trait, subset) %in% paste(diag_df$trait, diag_df$index)) %>% 
  ggplot() + geom_point(aes(x = h_sq, y = max_r2)) + 
  facet_grid(model~population, scales = 'free') 
  
```


# Comparing naive PTRS and ldblock based PTRS

## Overview

```{r plot1, fig.height=8, fig.width=6}
df_ld %>% dcast(population + ptrs_col + subset + trait + model ~ type, value.var = 'r2') %>% ggplot() + geom_point(aes(x = naive, y = ldblock, color = trait)) + facet_grid(population~model, scales = 'free') + geom_abline(slope = 1, intercept = 0)
```

## Best with each model

```{r best-british-vs, fig.height=8, fig.width=6}
df_ld_best_each = df_ld %>% group_by(population, subset, trait, model, type) %>% summarize(max_r2 = max(r2)) %>% ungroup()
df_ld_best_each %>% dcast(population + subset + trait + model ~ type, value.var = 'max_r2') %>% ggplot() + geom_point(aes(x = naive, y = ldblock, color = trait)) + facet_grid(population ~ model) + geom_abline(slope = 1, intercept = 0)
```

```{r vs-pve1}
df_ld_best_each$pop = df_ld_best_each$population
df_ld_best_each$pop[df_ld_best_each$pop %in% c('British-test', 'British-validation')] = 'British'
df_ld_best_each = inner_join(df_ld_best_each, regu, by = c('trait', 'pop' = 'population', 'model'))
```

```{r vs-pve-plot}
df_ld_best_each %>% 
  ggplot() + geom_point(aes(x = h_sq, y = max_r2, color = type), alpha = .3) + 
  facet_grid(model~population, scales = 'free') 
```