---
title: "Population assignment"
---

```{r setup}
rm(list = ls())
library(dplyr)
library(data.table)
library(pander)
library(ggplot2)
source('../code/rlib_doc.R')
```

# About 

The final query YAML is a bit different from `../output/query_martin_et_al_qtraits.yaml`. 
The query input and output is at `/vol/bmd/meliao/ukbiobank_query_2.tar` on `nucleus` now.

# Load data table

```{r load}
dat = fread('~/Desktop/ukbiobank_query_2/query_2_out.csv', data.table = F)
data_coding_for_ethnicity = read.delim2('~/Downloads/coding1001.tsv')
target_ethnicity_groups = c('Chinese', 'British', 'Indian', 'African')
data_coding_for_ethnicity = data_coding_for_ethnicity %>% filter(meaning %in% target_ethnicity_groups)
ethnicity_label = dat %>% select(ethnicity_x_instance_0_x_array_0, ethnicity_x_instance_1_x_array_0, ethnicity_x_instance_2_x_array_0) 
dat$ethnicity_agg = aggregate_instances(ethnicity_label)
dat = dat %>% filter(!is.na(ethnicity_agg)) %>% filter(ethnicity_agg %in% data_coding_for_ethnicity$coding)
dat = dat %>% inner_join(data_coding_for_ethnicity %>% select(coding, meaning), by = c('ethnicity_agg' = 'coding'))
dat %>% group_by(meaning) %>% summarize(nindiv = n()) %>% pander

pc_mat = dat[, colnames(dat) %in% paste0('pc', 1:10)]
rownames(pc_mat) = dat$eid
```

# Simple visualization

```{r vis, fig.height=10, fig.width=10}
vis = dat %>% select(pc1, pc2, pc3, pc4, pc5, pc6, pc7, pc8, pc9, pc10, meaning)
subsample = subsample_for_vis(vis$meaning, nmax = 1000)
p = myggpairs(vis[subsample, ] %>% select(-meaning), vis$meaning[subsample], alpha = .4)
p
```

# Run kmeans with everything

```{r, fig.height=10, fig.width=10}
kmean_assign = as.character(kmeans(vis %>% select(-meaning), centers = 5)$cluster)
p = myggpairs(vis[subsample, ] %>% select(-meaning), kmean_assign[subsample], alpha = .4)
p
```

# Do a post-filtering, how to set the cutoff?

Find the center of each cluster and treat it as a multivariate normal and calculate log-likelihood.

```{r filter}
ll_list = list()
for(i in unique(vis$meaning)) {
  sub = vis %>% filter(meaning == i) %>% select(-meaning)
  mean_hat = colMeans(sub)
  cov_hat = cov(sub)
  tmp = data.frame(ll = mvtnorm::dmvnorm(vis %>% select(-meaning), mean = mean_hat, sigma = cov_hat, log = T))
  colnames(tmp) = paste(i, 'dist')
  ll_list[[length(ll_list) + 1]] = tmp
}
df_ll = do.call(cbind, ll_list)
df_ll = df_ll %>% mutate(meaning = vis$meaning, eid = dat$eid)
df_ll %>% filter(subsample) %>% melt(id.vars = c('meaning', 'eid')) %>% ggplot() + geom_boxplot(aes(x = variable, y = value, color = meaning))
df_ll %>% head %>% pander
```

Get loglik in matched and un-match cases and see how loglik distribute.

```{r more}
df_ll_flat = df_ll[subsample, ] %>% melt(id.vars = c('meaning', 'eid'))
df_ll_flat$meaning = as.character(df_ll_flat$meaning)
df_ll_flat = df_ll_flat %>% mutate(is_match = meaning == stringr::str_remove(variable, ' dist'))
df_ll_flat %>% ggplot() + geom_jitter(aes(x = variable, y = value, color = is_match, group = is_match), height = 0, width = 0.3, alpha = 0.3) + coord_cartesian(ylim = c(-500, 0))
```

**Note**: the goal is to remove the 'outlier's which get mixed with un-matched individuals. 
**From this plot, we hard code the loglik threshold to -50

# Perform filtering

```{r do}
ll_cutoff = -50
eid_pass_cutoff = c()
for(pop in unique(df_ll$meaning)) {
  dist = paste(pop, 'dist')
  sub = df_ll %>% filter(meaning == pop)
  eid_pass_cutoff = c(eid_pass_cutoff, sub$eid[sub[, dist] > ll_cutoff])
}
dat_pass_cutoff = dat %>% filter(eid %in% eid_pass_cutoff)
```

# Plot indiv passing cutoff in PC space

```{r p, fig.height=10, fig.width=10}
vis = dat_pass_cutoff %>% select(pc1, pc2, pc3, pc4, pc5, pc6, pc7, pc8, pc9, pc10, meaning)
subsample = subsample_for_vis(vis$meaning, nmax = 1000)
p = myggpairs(vis[subsample, ] %>% select(-meaning), vis$meaning[subsample], alpha = .4)
p
```

# Save the cleaned-up data

```{r save}
saveRDS(list(dat_pass_pop_qc = dat_pass_cutoff, dat_ll = df_ll), '../output/query_first_attempt_with_population_qc.rds')
```
