---
title: "Query potentical technical covariates of monocyte count"
---

```{r setup}
rm(list = ls())
library(dplyr)
library(data.table)
library(ggplot2)
source('../code/rlib_doc.R')
```

# About 

It has drawn my attention that monocyte count distribution is a bit skewed and there are multiple modes. 
To look into this, I query several fields related to monocyte count.

# Load data table

```{r load}
dict = read.csv("http://biobank.ctsu.ox.ac.uk/~bbdatan/Data_Dictionary_Showcase.csv")
related_ids = data.frame(
  id = c(30134, 30132, 30133, 30131, 30130),
  name = c('acquisition_route', 'acquisition_time', 'device_id', 'freeze_cycle', 'monocyte')
)
myextract = dict %>% filter(FieldID %in% related_ids$id)
myextract = inner_join(myextract, related_ids, by = c('FieldID' = 'id'))
mydata = meta_info_to_query(myextract$name, myextract$FieldID, myextract$Instances, myextract$Array)
```

# Read in current YAML

```{r ryaml}
current_yaml = yaml::read_yaml('../misc/second_query.yaml')
```

Mainly to define individual filtering.

# Check whether the field exists in current ukbREST instance

```{r check}
good_fields = readRDS('../output/fields_on_nucleus.rds')
for(n in names(mydata)) {
  if(! mydata[[n]] %in% good_fields) {
    mydata[[n]] = NULL
  }
}
```


# Add the new entries to current YAML

```{r yaml}
current_yaml$data = mydata
yaml::write_yaml(current_yaml, '../output/query_monocyte.yaml')
```

# Query in ubkREST (lab server)

```{bash}
if [[ ! -f ../output/query_monocyte_output.csv ]]
then
  curl -X POST \
    -H "Accept: text/csv" \
    -F file=@../output/query_monocyte.yaml \
    -F section=data \
    http://127.0.0.1:12345/ukbrest/api/v1.0/query > ../output/query_monocyte_output.csv
fi
```

# And some quick exploration

Load query.

```{r lq}
df = fread('../output/query_monocyte_output.csv', header = T, sep = ',', data.table = F)
```


## Instance 0

```{r i1}
df_sub = df %>% select(eid, monocyte_x_instance_0_x_array_0, acquisition_time_x_instance_0_x_array_0, device_id_x_instance_0_x_array_0, acquisition_route_x_instance_0_x_array_0) %>% filter(!is.na(monocyte_x_instance_0_x_array_0))
df_sub = df_sub %>% mutate(acquisition_time_x_instance_0_x_array_0 = GENEAread::parse.time(acquisition_time_x_instance_0_x_array_0))
meta = list()
for(i in colnames(df_sub)) {
  if(i == 'eid') {
    next
  }
  meta[[length(meta) + 1]] = data.frame(var = i, nlevel = length(unique(df_sub[, i])))
  message('Number of unique values in ', i, ': ', length(unique(df_sub[, i])))
}
meta = do.call(rbind, meta)
subsample = subsample_for_vis(rep(1, nrow(df_sub)), nmax = 10000)

df_sub %>% filter(subsample) %>% ggplot() + geom_density(aes(x = monocyte_x_instance_0_x_array_0)) + coord_cartesian(xlim = c(0, 2))

df_plot = df_sub %>% filter(subsample)
meta$var = as.character(meta$var)
for(i in 1 : nrow(meta)) {
  if(meta$var[i] == 'monocyte_x_instance_0_x_array_0') {
    next
  }
  tmp = df_plot[, c(meta$var[i], 'monocyte_x_instance_0_x_array_0')]
  colnames(tmp)[1] = 'x'
  if(meta$nlevel[i] < 10) {
    p = ggplot(tmp) + geom_density(aes(x = monocyte_x_instance_0_x_array_0, fill = factor(x)), alpha = .5) + coord_cartesian(xlim = c(0, 2))
  } else {
    p = ggplot(tmp) + geom_point(aes(x = x, y = monocyte_x_instance_0_x_array_0), alpha = .5)
  }
  print(p + ggtitle(meta$var[i]))
}

ggplot(df_plot) + geom_boxplot(aes(x = factor(acquisition_route_x_instance_0_x_array_0), y = acquisition_time_x_instance_0_x_array_0))
```

## Instance 1

```{r i2}
df_sub = df %>% select(eid, monocyte_x_instance_1_x_array_0, acquisition_time_x_instance_1_x_array_0, device_id_x_instance_1_x_array_0, acquisition_route_x_instance_1_x_array_0) %>% filter(!is.na(monocyte_x_instance_1_x_array_0))
df_sub = df_sub %>% mutate(acquisition_time_x_instance_1_x_array_0 = GENEAread::parse.time(acquisition_time_x_instance_1_x_array_0))
meta = list()
for(i in colnames(df_sub)) {
  if(i == 'eid') {
    next
  }
  meta[[length(meta) + 1]] = data.frame(var = i, nlevel = length(unique(df_sub[, i])))
  message('Number of unique values in ', i, ': ', length(unique(df_sub[, i])))
}
meta = do.call(rbind, meta)
subsample = subsample_for_vis(rep(1, nrow(df_sub)), nmax = 10000)

df_sub %>% filter(subsample) %>% ggplot() + geom_density(aes(x = monocyte_x_instance_1_x_array_0)) + coord_cartesian(xlim = c(0, 2))

df_plot = df_sub %>% filter(subsample)
meta$var = as.character(meta$var)
for(i in 1 : nrow(meta)) {
  if(meta$var[i] == 'monocyte_x_instance_1_x_array_0') {
    next
  }
  tmp = df_plot[, c(meta$var[i], 'monocyte_x_instance_1_x_array_0')]
  colnames(tmp)[1] = 'x'
  if(meta$nlevel[i] < 10) {
    p = ggplot(tmp) + geom_density(aes(x = monocyte_x_instance_1_x_array_0, fill = factor(x)), alpha = .5) + coord_cartesian(xlim = c(0, 2))
  } else {
    p = ggplot(tmp) + geom_point(aes(x = x, y = monocyte_x_instance_1_x_array_0), alpha = .5)
  }
  print(p + ggtitle(meta$var[i]))
}

ggplot(df_plot) + geom_boxplot(aes(x = factor(acquisition_route_x_instance_1_x_array_0), y = acquisition_time_x_instance_1_x_array_0))
```

## Instance 3


```{r i3}
df_sub = df %>% select(eid, monocyte_x_instance_2_x_array_0, acquisition_time_x_instance_2_x_array_0, device_id_x_instance_2_x_array_0, acquisition_route_x_instance_2_x_array_0) %>% filter(!is.na(monocyte_x_instance_2_x_array_0))
df_sub = df_sub %>% mutate(acquisition_time_x_instance_2_x_array_0 = GENEAread::parse.time(acquisition_time_x_instance_2_x_array_0))
meta = list()
for(i in colnames(df_sub)) {
  if(i == 'eid') {
    next
  }
  meta[[length(meta) + 1]] = data.frame(var = i, nlevel = length(unique(df_sub[, i])))
  message('Number of unique values in ', i, ': ', length(unique(df_sub[, i])))
}
meta = do.call(rbind, meta)
subsample = subsample_for_vis(rep(1, nrow(df_sub)), nmax = 10000)

df_sub %>% filter(subsample) %>% ggplot() + geom_density(aes(x = monocyte_x_instance_2_x_array_0)) + coord_cartesian(xlim = c(0, 2))

df_plot = df_sub %>% filter(subsample)
meta$var = as.character(meta$var)
for(i in 1 : nrow(meta)) {
  if(meta$var[i] == 'monocyte_x_instance_2_x_array_0') {
    next
  }
  tmp = df_plot[, c(meta$var[i], 'monocyte_x_instance_2_x_array_0')]
  colnames(tmp)[1] = 'x'
  if(meta$nlevel[i] < 10) {
    p = ggplot(tmp) + geom_density(aes(x = monocyte_x_instance_2_x_array_0, fill = factor(x)), alpha = .5) + coord_cartesian(xlim = c(0, 2))
  } else {
    p = ggplot(tmp) + geom_point(aes(x = x, y = monocyte_x_instance_2_x_array_0), alpha = .5)
  }
  print(p + ggtitle(meta$var[i]))
}

ggplot(df_plot) + geom_boxplot(aes(x = factor(acquisition_route_x_instance_2_x_array_0), y = acquisition_time_x_instance_2_x_array_0))
```